---
title: "pw8105_hw1_gw2442"
output: github_document
date: "2022-09-17"
---

```{r setup, include = FALSE}
library(palmerpenguins)
library(tidyverse)
```

# Problem 1
### Loading the dataset:
```{r dataset_load}
data("penguins", package = "palmerpenguins")
```

### Names and values of important important variables in dataset "penguins":
```{r dataset_summary}
summary (penguins)
```
### Size of the dataset:
```{r dataset_size}
nrow(penguins)
ncol(penguins)
```
The dataset has 344 rows and 8 columns.

### Mean flipper length:
```{r flipper_mean}
data("penguins", package = "palmerpenguins")
mean(penguins$flipper_length_mm, na.rm = TRUE)
```
The mean flipper length is 200.92 mm. 

### Scatterplot of `flipper_length_mm` (y) vs `bill_length_mm` (x)
```{r yx_scatter}
ggplot(penguins, aes(x = bill_length_mm , y = flipper_length_mm, 
                     color = species)) + geom_point(na.rm = TRUE)
ggsave("penguins_scatter_plot.pdf")
```

# Problem 2
### Creating a dataframe 
```{r dataframe}
problem2_df = tibble(
  vec_numeric = rnorm(10, mean = 0 , sd = 1),
  vec_logical = vec_numeric > 0,
  vec_char = c("One", "Two", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten"),
  vec_factor = factor(c("1-3", "1-3", "1-3", "4-6", "4-6", "4-6", "7-10", "7-10", "7-10", "7-10"))
)
```
### Taking the mean of each variable in dataframe `problem2_df`
```{r dataframe_mean}
mean(problem2_df$vec_numeric)
mean(problem2_df$vec_logical)
mean(problem2_df$vec_char)
mean(problem2_df$vec_factor)
```
There is an output for the variables `vec_numeric`  and `vec_logical` when calculating the mean. There was no mean output for varaibles `vec_char` and `vec_factor`.  

### Converting variables to numeric variables
```{r dataframe_numeric, results = 'hide', warning = FALSE}
as.numeric(problem2_df$vec_logical)
as.numeric(problem2_df$vec_char)
as.numeric(problem2_df$vec_factor)
```
The `as.numeric` function produces an numeric output for `vec_logical` comprised of 0's and 1's. 
The `as.numeric` function does not produce an output for `vec_char`. It outputs NA.
The `as.numeric` function produces an numeric output for `vec_factor` comprised of 1's, 2's, and 3's. 


When the `as.numeric` function is applied to the logic variable `vec_logical`, the outputs for the variable are either 1 (corresponding to TRUE) or 0 (corresponding to FALSE). This explains the output mean for `vec_logical`. When the mean is calculated for `vec_logical`, R takes the TRUE and FALSE values in the dataframe and corresponds it to 1 or 0, and then calculates the mean value, resulting in a valid output when running `mean(problem2_df$vec_logical)`. 

When the `as.numeric` function is applied to the factor variable `vec_factor`, the outputs for the variable are either 1, 2, or 3. These three numbers correspond to the three factor "levels" specified the dataframe. When calculating the mean for `vec_factor`, the factor "levels" were specified as characters, and therefore a mean could not be calculated. However, applying the `as.numeric` function converts the factor "levels" into numeric "levels", resulting in the outputs of 1, 2, or 3. 

The character variable `vec_char`, on the other hand, is comprised of letters and has no numeric value, unlike the logic and factor variables. When the `as.numeric` function is applied to the character variable, it is unable to convert the characters into a numeric variable, resulting in the NA output.

